---
title: "MAG - 'Climatology'"
author: "overgoor"
params:
  date: "Sys.Date()"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    code_folding: hide
    self_contained: yes
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo=F, warning=F, cache=T, message=F, fig.width=6, fig.height=3)
suppressPackageStartupMessages(library('ggplot2'))
ggplot2::theme_set(theme_minimal())
library(tidyr)
library(dplyr)
library(readr)
library(mlogit)
library(stringr)
```

```{r}
# cleaner theme
my_theme <- function(base_size=10) {
  # Set the base size
  theme_bw(base_size=base_size) +
    theme(
      # Center title
      plot.title = element_text(hjust = 0.5),
      # Make the background white
      panel.background=element_rect(fill='white', colour='white'),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      # Minimize margins
      plot.margin=unit(c(0.2, 0.2, 0.2, 0.2), "cm"),
      panel.margin=unit(0.25, "lines"),
      # Tiny space between axis labels and tick labels
      axis.title.x=element_text(margin=ggplot2::margin(t=6.0)),
      axis.title.y=element_text(margin=ggplot2::margin(r=6.0)),
      # Simplify the legend
      legend.key=element_blank(),
      legend.title=element_blank(),
      legend.background=element_rect(fill='transparent')
    )
}
```

# Read Data

```{r data}
d_raw = read_csv("~/choosing_to_grow/data_academic/processed/mag_cli.txt", col_types='cccddccc')
# explode citations into edges
edges = d_raw %>%
  separate_rows(references, sep=',') %>%
  select(id, cites=references, year_cited=year) %>%
  left_join(d_raw %>% select(cites=id, year_published=year))
```

# Plots

## Role of Year

```{r plot_dist_time1}
# distribution of when papers are published
d_raw %>%
  group_by(year) %>% summarize(n=n()) %>%
  ggplot(aes(year, n)) + geom_line() +
    scale_x_continuous("Year", limits=c(1980,2020)) +
    scale_y_continuous("# papers") +
    ggtitle("Number of papers per year") +
    my_theme()
```

_ _ 

```{r plot_dist_time2}
# distribution of time between a papers publishing year and year of citation
edges %>%
  mutate(delta=year_cited-year_published) %>% 
  mutate(delta=ifelse(delta < 0, NA, ifelse(delta > 40, 40, delta))) %>%
  group_by(delta) %>% summarize(n=n()) %>%
  filter(!is.na(delta)) %>%
  ggplot(aes(delta, n)) + geom_line() +
    scale_x_continuous("Delta (years)") +
    scale_y_continuous("# papers") +
    ggtitle("Distribution of Years to citation") +
    my_theme()
```

## Share of citations observed

```{r data_coverage}
# how many of the citations in "n_citation" do we actually observe in the data?
DF = inner_join(d_raw,
           edges %>% group_by(cites) %>% summarize(n_citation_data=n()),
           by=c('id'='cites')) %>%
  select(id, n_citation, n_citation_data) %>%
  mutate(p_citation_data=n_citation_data/n_citation) %>%
  filter(!is.na(id))
```

```{r plot_coverage_scatter, eval=F}
# scatter
ggplot(DF, aes(n_citation, n_citation_data)) + geom_point(alpha=0.2) +
  scale_x_log10("(log) # citations (stated)") +
  scale_y_log10("# citations (observed)") +
  geom_line(data=data.frame(x=10:10000), aes(x, x), color='blue') + 
  ggtitle("Citations - Stated vs Observed (individual)") +
  my_theme()
```

```{r plot_coverage_group}
# grouped
DF %>% group_by(n_citation) %>% summarize(stat=mean(n_citation_data)) %>%
  ggplot(aes(n_citation, stat)) + geom_line() + #geom_point(alpha=0.2) +
    scale_x_log10("(log) # citations (stated)") +
    scale_y_log10("# citations (observed)") +
    geom_line(data=data.frame(x=10:10000), aes(x, x), color='blue') + 
    ggtitle("Citations - Stated vs Observed (grouped)") +
    my_theme()
```

```{r plot_coverage_density}
# density
ggplot(DF, aes(p_citation_data)) + geom_density() +
  scale_x_continuous("Share of citations") +
  ggtitle("Share of stated citations observed") +
  my_theme()
```


## Degree distribution

```{r plot_degdist}
# degree distribution
rbind(
    DF %>% mutate(stat=n_citation) %>% filter(stat > 50) %>% group_by(stat) %>% summarize(n=n()) %>% mutate(g='Stated'),
    DF %>% mutate(stat=n_citation_data) %>% group_by(stat) %>% summarize(n=n()) %>% mutate(g='Observed')
  ) %>%
  ggplot(aes(stat, n)) + geom_point() + 
    scale_x_log10("(log) # citations") +
    scale_y_log10("(log) # papers") +
    ggtitle("Degree Distribution") +
    facet_wrap(~g) +
    my_theme()
```


## Author distribution

```{r plot_authdist, fig.width=4, fig.height=3}
d_raw %>% select(authors) %>%
  separate_rows(authors, sep=',') %>%
  group_by(authors) %>% summarize(n_papers=n()) %>%
  group_by(n_papers) %>% summarize(n=n()) %>%
  ggplot(aes(n_papers, n)) + geom_point() +
    ggtitle("Distribution of number of papers/author") +
    scale_x_log10("(log) Number of papers") +
    scale_y_log10("(log) Number of Authors") +
    my_theme()
```

## Keyword distribution

```{r plot_keydist}
d_raw %>% select(keywords) %>%
  separate_rows(keywords, sep=',') %>%
  mutate(keyword=keywords) %>% filter(!is.na(keyword)) %>%
  group_by(keywords) %>% summarize(n=n()) %>%
  arrange(-n) %>% head(n=10) %>%
  kable(format='markdown')
```


## Individual paper's timeline

```{r plot_timeline}
# timeline of citations
edges %>% 
  filter(cites=='9c024621-3fbb-411d-ae8c-984585f98221') %>%
  group_by(year_cited) %>% summarize(n=n()) %>%
  arrange(year_cited) %>%
  ggplot(aes(year_cited, n)) + geom_point() + geom_line() +
    ggtitle("Citations for 9c024621-3fbb-411d-ae8c-984585f98221") +
    xlab("Year") + ylab("Number of citations") +
    my_theme()
```


# Model

Data construction process:

* sample 1000 citations
* for each actual citation, sample 24 non-cited papers (from before publication date)
* for each of the (paper,option) pairs, compute features (n citations, years since, has same author)


```{r model_data_1}
set.seed(100)
## sample actual choices
dm1 <- edges %>%
  # only look at citations within dataset
  filter(!is.na(year_published), year_published < year_cited, year_cited==2014) %>%
  sample_n(1000) %>%
  mutate(y=1) %>%
  select(paper_id=id, paper_year=year_cited, option_id=cites, option_year=year_published, y)
```


```{r model_data_2}
## sample reduced choice sets
dm2 <- dm1 %>%
  mutate(t='x') %>%
  select(paper_id, paper_year, t) %>%
  # cross join with all possible choices to create full choice set
  full_join(d_raw %>%
              filter(year > 1930) %>%
              select(option_id=id, option_year=year) %>%
              #sample_n(1000) %>%
              mutate(t='x'), by='t') %>%
  filter(option_year < paper_year) %>%
  # sample only 20 choices for reduced choice set
  group_by(paper_id) %>% sample_n(24) %>% ungroup() %>%
  mutate(y=0) %>% select(paper_id, paper_year, option_id, option_year, y) %>%
  # make sure they weren't actually cited..
  left_join(edges, by=c("paper_id"="id","option_id"="cites")) %>%
  filter(is.na(year_cited)) %>% select(-year_cited, -year_published)
```


```{r model_data_3}
# helper function to count number of overlapping items in comma-separated list string
str_overlap <- Vectorize(function(s1, s2) {
  #l1 = lapply(str_split(s1, ',')[[1]], function(x) { y = str_split(tolower(x), ' ')[[1]]; sprintf('%s_%s', str_sub(y[1], 1, 1), y[length(y)]) })
  #l2 = lapply(str_split(s2, ',')[[1]], function(x) { y = str_split(tolower(x), ' ')[[1]]; sprintf('%s_%s', str_sub(y[1], 1, 1), y[length(y)]) })
  #intersect(l1, l2) %>% length() > 0
  intersect(str_split(s1, ',')[[1]], str_split(s2, ',')[[1]]) %>% length()
})

dm3 <- rbind(dm1, dm2)

## Compute features
DM <- dm3 %>%
  # n_citations
  left_join(
    # compute number of citations **at time of choosing papers' publishing**
    inner_join(dm3, edges %>% select(cites, year_cited), by=c('option_id'='cites')) %>%
      filter(paper_year > year_cited) %>%
      group_by(paper_id, option_id) %>% summarize(n_citations=n()),
    by=c('paper_id','option_id')
  ) %>%
  mutate(n_citations=ifelse(is.na(n_citations), 0, n_citations)) %>%
  # years between
  mutate(delta_years = paper_year - option_year) %>%
  # overlapping authors
  left_join(
    dm3 %>% select(paper_id, option_id) %>%
      left_join(d_raw %>% select(id, a=authors), by=c('paper_id' ='id')) %>%
      left_join(d_raw %>% select(id, b=authors), by=c('option_id'='id')) %>%
      mutate(has_same_author=ifelse(str_overlap(a, b) > 0, 1, 0)) %>% select(-a, -b),
    by=c('paper_id','option_id')
  ) %>%
  # overlapping keywords
  left_join(
    dm3 %>% select(paper_id, option_id) %>%
      left_join(d_raw %>% select(id, a=keywords), by=c('paper_id' ='id')) %>%
      left_join(d_raw %>% select(id, b=keywords), by=c('option_id'='id')) %>%
      mutate(n_same_keywords=str_overlap(a, b)) %>% select(-a, -b),
    by=c('paper_id','option_id')
  )

H <- DM %>% 
  group_by(paper_id) %>% mutate(alt_id=row_number()) %>% ungroup() %>%
  select(paper_id, alt_id, y, n_citations, delta_years, has_same_author, n_same_keywords) %>% as.data.frame() %>%
  mlogit.data(shape="long", chid.var='paper_id', choice="y", alt.var='alt_id', varying=c(4:7))
```


```{r model_model}
stargazer::stargazer(
  #mlogit(y ~ n_citations | 0, H),
  mlogit(y ~ log(n_citations+1) | 0, H),
  #mlogit(y ~ log(n_citations+1) + delta_years | 0, H),
  mlogit(y ~ log(n_citations+1) + log(delta_years) | 0, H),
  mlogit(y ~ log(n_citations+1) + log(delta_years) + has_same_author | 0, H),
  mlogit(y ~ log(n_citations+1) + log(delta_years) + has_same_author + n_same_keywords| 0, H),  
  type='text')
```
